### 2016-04-27 Background/php.tex

\lstset{language=PHP,caption={PHP program with e-mail header injection vulnerability.},label={code:phpemi}}
\begin{lstlisting}
$from = $_REQUEST['email'];
$subject = "Hello Sai Pc";
$message = "We need you to reset your password";
$to = "schand31@asu.edu";

// example attack string to be injected as the value for
// $_REQUEST['email'] => 'sai@sai.com\nCC:spc@spc.com'
$retValue = mail($to, $subject, $message, "From: $from");
// E-Mail gets sent to both schand31@asu.edu AND spc@spc.com
\end{lstlisting}