### 2016-04-27 Experiment/platform.tex

\section[Platform]{Platforms and Software}

We enumerate the platforms and the software used for our project in Table~\ref{tab:platsw}.
\begin{table}[!htbp]
	\centering
	\begin{tabular}{|c|c|}
		\hline
		Operating system & Ubuntu 14.04\\
		\hline
		Server & Apache - 2.4.17\\
		\hline
		Database & MariaDB - 10.1.9\\
		\hline
		Mail Server & Postfix - 2.11.0\\
		\hline
		Other software used & Mailcatcher, PostMan, HTTPRequester, RabbitMQ\\
		\hline
	\end{tabular}
	\caption[\titlecap{Platforms and software}]{Platforms and software used for our project.}
	\label{tab:platsw}
\end{table}