### 2016-04-27 Results/fuzz_result.tex

\section{Fuzzed data}
We performed our fuzzing attempts on the gathered data with both the regular payload and malicious payload. Table~\ref{tab:fuzzed_data} shows the quantity of e-mails that we received for each payload. We explain in detail what each piece of data shown in the table represents, in the following sections.
\input{Results/fuzzed_data}
\paragraph{E-Mail received from forms}
The e-mails that we received can be broadly categorized into two categories:
\begin{enumerate}
	\item E-Mails due to regular payload\\
	This represents the total number of websites that sent e-mails to us. This indicates that we were able to successfully submit the forms on these sites.

	\item E-Mails due to malicious payload\\
    Once we receive an e-mail from a website due to the regular payload, we go back and fuzz those forms with more malicious payloads. This field, in essence, represents the total number of unique URLs that contain E-Mail Header Injection vulnerability.
\end{enumerate}