### 2016-04-27 Results/pre_result.tex

\section{Collected Data}
From our extensive crawl of the web, we were able to gather the data shown in Table~\ref{tab:data}. The following paragraphs describe in detail what each kind of data represents, and what they signify.

\input{data}

\paragraph{URL's crawled}
This represents the total number of unique URLs that we crawled on the World Wide Web. It is to be noted that this quantity refers to unique URL's, and not websites.

\paragraph{Total Forms found}
This represents the total number of forms that were found on the URL's crawled. We found a total of \forms\ forms from 
\uniqueforms\ unique domains.

\paragraph{Forms with E-Mail Fields}
This represents the total number of forms that were found to contain e-mail fields. We found \emailforms\ such forms from \uniqueemailforms\ unique domains.