### 2016-04-27 data_analysis.tex

\begin{table}[!htbp]
	\centering
	\begin{tabular}{|c|l|c|}
		\hline
		\multicolumn{1}{|c|}{\textbf{S.No}} &
		\multicolumn{1}{c|}{\textbf{Type of Injection}} &
		\multicolumn{1}{p{3cm}|}{\centering \textbf{No. of e-mails received}}\\
		\hline
		1 & E-Mail Header Injections with `bcc' header & 180\\
		\hline
		2 & E-Mail Header Injections with `x-check' header & 153\\
		\hline
		3 & `To header' injections alone & 83\\
		\hline
		4 & E-Mail Header Injections with `bcc' and `x-check' headers & 136\\
		\hline
		5 & Both `To header' injections and x-check headers & 9\\
		\hline
		6 & `x-check' headers found in `nuser' e-mails & 40\\
		\hline
		7 & Unique `x-check' headers found in `nuser' e-mails & 16\\
		\hline
	\end{tabular}
	\caption[\titlecap{Analysis of the data}]{Classification of the e-mails that we received into broad categories of the vulnerability.}
	\label{tab:analysis}
\end{table}