### 2016-04-27 dis.tex

\documentclass[12pt,letterpaper]{report}
\usepackage[square,numbers]{natbib}
\usepackage{geometry}
\usepackage{fancyhdr}
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsbsy}
\usepackage{dcolumn,array}
\usepackage{tocloft}

% for title case
\usepackage{titlecaps}
\Addlcwords{and, the, or, of, that, our, by}

% avoid widows
\usepackage[all]{nowidow}
\clubpenalty=10000

\usepackage{asudis}
%% These are for including code
%% and for upquotes
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{listingsutf8}
\usepackage{comment}
\usepackage{color}
\usepackage{dirtytalk} % for quotations
\usepackage[justification=centering]{caption}
\usepackage[pageanchor=true,plainpages=false,pdfpagelabels,bookmarks,bookmarksnumbered,hidelinks]{hyperref}

% for footnotes in tables
\usepackage{tablefootnote}

% for arithmetic
\usepackage{xintexpr}

\usepackage{textcomp}     % access \textquotesingle

\definecolor{mygray}{rgb}{0.6,0.6,0.6}
\definecolor{lightgray}{rgb}{0.92,0.92,0.92}
\definecolor{darkgreen}{rgb}{0,0.7,0}
% default options for listings
\lstset{
	backgroundcolor=\color{lightgray},
	basicstyle=\footnotesize,
	breaklines=true, 
	captionpos=b,
	commentstyle=\color{darkgreen}, 
	frame=single,
	keywordstyle=\color{blue}, 
	numbers=left,
	numbersep=5pt,
	numberstyle=\tiny\color{mygray},
	rulecolor=\color{black},
	showstringspaces=false,
	upquote=true
}

\newcommand{\dq}[1]{``{#1}''}
%% THIS IS THE DATA FOR OUR THESIS, UPDATING HERE WILL UPDATE EVERYWHERE %%
\newcommand{\urls}{17,686,161}

\newcommand{\forms}{5,562,533}
\newcommand{\formsDelta}{31.45\%}

\newcommand{\emailforms}{918,988}
\newcommand{\emailformsDelta}{16.52\%}

\newcommand{\fuzzed}{756,588}
\newcommand{\fuzzedDelta}{82.33\%}

\newcommand{\recd}{44,735}
\newcommand{\recdDelta}{5.91\%}

\newcommand{\malfuzzed}{38,989}
\newcommand{\malfuzzedDelta}{87.16\%}

\newcommand{\success}{433}
\newcommand{\successDelta}{1.11\%}

\newcommand{\domains}{182}

% these refer to unique domains, not unique forms
\newcommand{\uniqueforms}{650,394}
\newcommand{\uniqueemailforms}{151,795}

\begin{document}
%-----------------------front matter
\pagenumbering{roman}
\title{E-Mail Header Injections\\
	An Analysis of the World Wide Web}

\author{Sai Prashanth Chandramouli}
\degreeName{Master of Science}
\paperType{Thesis}
\defensemonth{April}
\defenseyear{2016}
\gradmonth{May}
\gradyear{2016}
\chair{Adam Doup\'{e}}
\memberOne{Gail-Joon Ahn}
\memberTwo{Ziming Zhao}
%%\memberThree{Your third member}
%%\memberFour{Your third member}

\maketitle
\doublespace
\include{abstract}
\dedicationpage{To my mother and father, for giving me the life I dreamt of,\\
To my sister, who constantly made me do better just to keep up with her,\\
To my family in Phoenix, for always being there,\\
To God, for making me so lucky, for letting me be strong when I had nothing, and making me believe when no one else would have.}
\begin{acknowledgements}
	\input{sample_acknowledgements}
\end{acknowledgements}
\tableofcontents
% This puts the word "Page" right justified above everything else.
\addtocontents{toc}{~\hfill Page\par}
% Asking LaTeX for a new page here guarantees that the LOF is on a separate page
% after the TOC ends.
\newpage
% Making the LOT and LOF "parts" rather than chapters gets them indented at
% level -1 according to the chart: top of page 4 of the document at
% ftp://tug.ctan.org/pub/tex-archive/macros/latex/contrib/tocloft/tocloft.pdf
\addcontentsline{toc}{part}{LIST OF TABLES}
\renewcommand{\cftlabel}{Table}
\listoftables
% This gets the headers for the LOT right on the first page.  Subsequent pages
% are handled by the fancyhdr code in the asudis.sty file.
\addtocontents{lot}{Table~\hfill Page \par}
\newpage
\addcontentsline{toc}{part}{LIST OF FIGURES}
\addtocontents{toc}{CHAPTER \par}
\renewcommand{\cftlabel}{Figure}
% This gets the headers for the LOF right on the first page.  Subsequent pages
% are handled by the fancyhdr code in the asudis.sty file.
\addtocontents{lof}{Figure~\hfill Page \par}
\listoffigures
%-----------------------body
\doublespace
\pagenumbering{arabic}
\include{introduction}
\include{problem_background}
\include{sys_design}
\include{exp_setup}
\include{results}
\include{discussion}
\include{rel_work}
\include{conclusion}

%-----------------------back matter
{\singlespace
% Making the references a "part" rather than a chapter gets it indented at
% level -1 according to the chart: top of page 4 of the document at
% ftp://tug.ctan.org/pub/tex-archive/macros/latex/contrib/tocloft/tocloft.pdf
\addcontentsline{toc}{part}{REFERENCES}
\bibliographystyle{asudis}
\bibliography{biblio}}
\nocite{*}
\renewcommand{\chaptername}{APPENDIX}
\addtocontents{toc}{APPENDIX \par}
\appendix
\include{appendix1}
\end{document}