### 2016-04-27 sample_acknowledgements.tex

A project of this size is never easy to complete without the help and support of other people. I would like to take this opportunity to thank some of them.

This thesis would not have been possible without the help and guidance of my thesis advisor, and committee chair --- the brilliant Dr.\,Adam Doup\'{e}.
This project was his brainchild, and he held my hand through the entire project. Thank you for everything you did, Adam.

I would like to thank Dr.\,Gail-Joon Ahn, for being part of the committee, for all his help, and his valuable input on the changes to be made to make the project more impactful.

I would also like to thank Dr.\,Ziming Zhao, for being a part of my committee, and for the constant motivation.

I would like to thank the members of the SEFCOM, for their support, and would like to especially thank Mike Mabey, for helping set up the infrastructure for this project, and Marthony Taguinod, for helping me document this project.