### 2016-05-06 paper.tex
lil adds

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate-05-2015}

\usepackage[square,numbers]{natbib}
\usepackage[justification=centering]{caption}
\usepackage[pageanchor=true,plainpages=false,pdfpagelabels,bookmarks,bookmarksnumbered,hidelinks]{hyperref}
\usepackage{color}
\definecolor{mygray}{rgb}{0.6,0.6,0.6}
\definecolor{lightgray}{rgb}{0.92,0.92,0.92}
\definecolor{darkgreen}{rgb}{0,0.7,0}

% for footnotes in tables
\usepackage{tablefootnote}

% for title caps
\usepackage{titlecaps}
\Addlcwords{and, the, or, of, that, our, by, a, prevent, for}

% default options for listings
\usepackage{listings}
\usepackage{listingsutf8}
\usepackage{textcomp}     % access \textquotesingle
\lstset{
	backgroundcolor=\color{lightgray},
	basicstyle=\footnotesize,
	breaklines=true, 
	captionpos=b,
	commentstyle=\color{darkgreen}, 
	frame=single,
	keywordstyle=\color{blue}, 
	numbers=left,
	numbersep=5pt,
	numberstyle=\tiny\color{mygray},
	rulecolor=\color{black},
	showstringspaces=false,
	upquote=true
}

\newcommand{\dq}[1]{``{#1}''}
%% THIS IS THE DATA FOR OUR THESIS, UPDATING HERE WILL UPDATE EVERYWHERE %%
\newcommand{\urls}{21,675,680}

\newcommand{\forms}{6,794,917}
\newcommand{\formsDelta}{31.35\%}

\newcommand{\emailforms}{1,132,157}
\newcommand{\emailformsDelta}{16.66\%}

\newcommand{\fuzzed}{934,016}
\newcommand{\fuzzedDelta}{82.50\%}

\newcommand{\recd}{52,724}
\newcommand{\recdDelta}{5.64\%}

\newcommand{\malfuzzed}{46,156}
\newcommand{\malfuzzedDelta}{87.54\%}

\newcommand{\success}{496}
\newcommand{\successDelta}{1.07\%}

\newcommand{\domains}{222}
\newcommand{\ips}{292}
% these refer to unique domains, not unique forms
\newcommand{\uniqueforms}{1,019,921}
\newcommand{\uniqueemailforms}{197,570}

\begin{document}

	% Copyright
	\setcopyright{acmcopyright}
	%\setcopyright{acmlicensed}
	%\setcopyright{rightsretained}
	%\setcopyright{usgov}
	%\setcopyright{usgovmixed}
	%\setcopyright{cagov}
	%\setcopyright{cagovmixed}

	% DOI
	\doi{10.475/123_4}

	% ISBN
	\isbn{123-4567-24-567/08/06}

	%Conference
	\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

	\acmPrice{\$15.00}

	%
	% --- Author Metadata here ---
	\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
	%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
	%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
	% --- End of Author Metadata ---

	\title{E-jection fraction - Tracking how your website pumps out E-Mails}
	%
	% You need the command \numberofauthors to handle the 'placement
	% and alignment' of the authors beneath the title.
	%
	% For aesthetic reasons, we recommend 'three authors at a time'
	% i.e. three 'name/affiliation blocks' be placed beneath the title.
	%
	% NOTE: You are NOT restricted in how many 'rows' of
	% "name/affiliations" may appear. We just ask that you restrict
	% the number of 'columns' to three.
	%
	% Because of the available 'opening page real-estate'
	% we ask you to refrain from putting more than six authors
	% (two rows with three columns) beneath the article title.
	% More than six makes the first-page appear very cluttered indeed.
	%
	% Use the \alignauthor commands to handle the names
	% and affiliations for an 'aesthetic maximum' of six authors.
	% Add names, affiliations, addresses for
	% the seventh etc. author(s) as the argument for the
	% \additionalauthors command.
	% These 'additional authors' will be output/set for you
	% without further effort on your part as the last section in
	% the body of your article BEFORE References or any Appendices.

	\numberofauthors{8} %  in this sample file, there are a *total*
	% of EIGHT authors. SIX appear on the 'first-page' (for formatting
	% reasons) and the remaining two appear in the \additionalauthors section.
	%
	\author{
		% You can go ahead and credit any number of authors here,
		% e.g. one 'row of three' or two rows (consisting of one row of three
		% and a second row of one, two or three).
		%
		% The command \alignauthor (no curly braces needed) should
		% precede each author name, affiliation/snail-mail address and
		% e-mail address. Additionally, tag each line of
		% affiliation/address with \affaddr, and tag the
		% e-mail address with \email.
		%
		% 1st. author
		\alignauthor
		Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
		\affaddr{Institute for Clarity in Documentation}\\
		\affaddr{1932 Wallamaloo Lane}\\
		\affaddr{Wallamaloo, New Zealand}\\
		\email{trovato@corporation.com}
		% 2nd. author
		\alignauthor
		G.K.M. Tobin\titlenote{The secretary disavows
			any knowledge of this author's actions.}\\
		\affaddr{Institute for Clarity in Documentation}\\
		\affaddr{P.O. Box 1212}\\
		\affaddr{Dublin, Ohio 43017-6221}\\
		\email{webmaster@marysville-ohio.com}
		% 3rd. author
		\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
			one who did all the really hard work.}\\
		\affaddr{The Th{\o}rv{\"a}ld Group}\\
		\affaddr{1 Th{\o}rv{\"a}ld Circle}\\
		\affaddr{Hekla, Iceland}\\
		\email{larst@affiliation.org}
		\and  % use '\and' if you need 'another row' of author names
		% 4th. author
		\alignauthor Lawrence P. Leipuner\\
		\affaddr{Brookhaven Laboratories}\\
		\affaddr{Brookhaven National Lab}\\
		\affaddr{P.O. Box 5000}\\
		\email{lleipuner@researchlabs.org}
		% 5th. author
		\alignauthor Sean Fogarty\\
		\affaddr{NASA Ames Research Center}\\
		\affaddr{Moffett Field}\\
		\affaddr{California 94035}\\
		\email{fogartys@amesres.org}
		% 6th. author
		\alignauthor Charles Palmer\\
		\affaddr{Palmer Research Laboratories}\\
		\affaddr{8600 Datapoint Drive}\\
		\affaddr{San Antonio, Texas 78229}\\
		\email{cpalmer@prl.com}
	}
	% There's nothing stopping you putting the seventh, eighth, etc.
	% author on the opening page (as the 'third row') but we ask,
	% for aesthetic reasons that you place these 'additional authors'
	% in the \additional authors block, viz.
	\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
		email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
		(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
	\date{30 July 1999}
	% Just remember to make sure that the TOTAL number of authors
	% is the number that will appear on the first page PLUS the
	% number that will appear in the \additionalauthors section.

	\maketitle
	\input{abstract}

	%
	% The code below should be generated by the tool at
	% http://dl.acm.org/ccs.cfm
	% Please copy and paste the code instead of the example below. 
	%
	\begin{CCSXML}
		<ccs2012>
		<concept>
		<concept_id>10010520.10010553.10010562</concept_id>
		<concept_desc>Computer systems organization~Embedded systems</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10010520.10010575.10010755</concept_id>
		<concept_desc>Computer systems organization~Redundancy</concept_desc>
		<concept_significance>300</concept_significance>
		</concept>
		<concept>
		<concept_id>10010520.10010553.10010554</concept_id>
		<concept_desc>Computer systems organization~Robotics</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		<concept>
		<concept_id>10003033.10003083.10003095</concept_id>
		<concept_desc>Networks~Network reliability</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		</ccs2012>  
	\end{CCSXML}

	\ccsdesc[500]{Computer systems organization~Embedded systems}
	\ccsdesc[300]{Computer systems organization~Redundancy}
	\ccsdesc{Computer systems organization~Robotics}
	\ccsdesc[100]{Networks~Network reliability}

	%
	% End generated code
	%

	%
	%  Use this command to print the description
	%
	\printccsdesc

	% We no longer use \terms command
	%\terms{Theory}

	\keywords{ACM proceedings; \LaTeX; text tagging}

	\input{introduction}

	\input{problem_background}

	\section{The {\secit Body} of The Paper}
	Typically, the body of a paper is organized
	into a hierarchical structure, with numbered or unnumbered
	headings for sections, subsections, sub-subsections, and even
	smaller sections.  The command \texttt{{\char'134}section} that
	precedes this paragraph is part of such a
	hierarchy.\footnote{This is the second footnote.  It
		starts a series of three footnotes that add nothing
		informational, but just give an idea of how footnotes work
		and look. It is a wordy one, just so you see
		how a longish one plays out.} \LaTeX\ handles the numbering
	and placement of these headings for you, when you use
	the appropriate heading commands around the titles
	of the headings.  If you want a sub-subsection or
	smaller part to be unnumbered in your output, simply append an
	asterisk to the command name.  Examples of both
	numbered and unnumbered headings will appear throughout the
	balance of this sample document.

	Because the entire article is contained in
	the \textbf{document} environment, you can indicate the
	start of a new paragraph with a blank line in your
	input file; that is why this sentence forms a separate paragraph.

	\subsection{Type Changes and {\subsecit Special} Characters}
	We have already seen several typeface changes in this sample.  You
	can indicate italicized words or phrases in your text with
	the command \texttt{{\char'134}textit}; emboldening with the
	command \texttt{{\char'134}textbf}
	and typewriter-style (for instance, for computer code) with
	\texttt{{\char'134}texttt}.  But remember, you do not
	have to indicate typestyle changes when such changes are
	part of the \textit{structural} elements of your
	article; for instance, the heading of this subsection will
	be in a sans serif\footnote{A third footnote, here.
		Let's make this a rather short one to
		see how it looks.} typeface, but that is handled by the
	document class file. Take care with the use
	of\footnote{A fourth, and last, footnote.}
	the curly braces in typeface changes; they mark
	the beginning and end of
	the text that is to be in the different typeface.

	You can use whatever symbols, accented characters, or
	non-English characters you need anywhere in your document;
	you can find a complete list of what is
	available in the \textit{\LaTeX\
		User's Guide}\cite{Lamport:LaTeX}.

	\subsection{Tables}
	Because tables cannot be split across pages, the best
	placement for them is typically the top of the page
	nearest their initial cite.  To
	ensure this proper ``floating'' placement of tables, use the
	environment \textbf{table} to enclose the table's contents and
	the table caption.  The contents of the table itself must go
	in the \textbf{tabular} environment, to
	be aligned properly in rows and columns, with the desired
	horizontal and vertical rules.  Again, detailed instructions
	on \textbf{tabular} material
	is found in the \textit{\LaTeX\ User's Guide}.

	Immediately following this sentence is the point at which
	Table 1 is included in the input file; compare the
	placement of the table here with the table in the printed
	dvi output of this document.

	\begin{table}
		\centering
		\caption{Frequency of Special Characters}
		\begin{tabular}{|c|c|l|} \hline
			Non-English or Math&Frequency&Comments\\ \hline
			\O & 1 in 1,000& For Swedish names\\ \hline
			$\pi$ & 1 in 5& Common in math\\ \hline
			\$ & 4 in 5 & Used in business\\ \hline
			$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
			\hline\end{tabular}
	\end{table}

	To set a wider table, which takes up the whole width of
	the page's live area, use the environment
	\textbf{table*} to enclose the table's contents and
	the table caption.  As with a single-column table, this wide
	table will ``float" to a location deemed more desirable.
	Immediately following this sentence is the point at which
	Table 2 is included in the input file; again, it is
	instructive to compare the placement of the
	table here with the table in the printed dvi
	output of this document.

	\begin{table*}
		\centering
		\caption{Some Typical Commands}
		\begin{tabular}{|c|c|l|} \hline
			Command&A Number&Comments\\ \hline
			\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
			\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
			\texttt{{\char'134}table}& 300 & For tables\\ \hline
			\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
	\end{table*}
	% end the environment with {table*}, NOTE not {table}!

	\subsection{Figures}
	Like tables, figures cannot be split across pages; the
	best placement for them
	is typically the top or the bottom of the page nearest
	their initial cite.  To ensure this proper ``floating'' placement
	of figures, use the environment
	\textbf{figure} to enclose the figure and its caption.

	This sample document contains examples of \textbf{.eps} files to be
	displayable with \LaTeX.  If you work with pdf\LaTeX, use files in the
	\textbf{.pdf} format.  Note that most modern \TeX\ system will convert
	\textbf{.eps} to \textbf{.pdf} for you on the fly.  More details on
	each of these is found in the \textit{Author's Guide}.

	\begin{figure}
		\centering
		%\includegraphics{System/fuzzer_design}
		\caption{A sample black and white graphic.}
	\end{figure}

	\begin{figure}
		\centering
		%\includegraphics[height=1in, width=1in]{fly.eps}
		\caption{A sample black and white graphic
			that has been resized with the \texttt{includegraphics} command.}
	\end{figure}

	As was the case with tables, you may want a figure
	that spans two columns.  To do this, and still to
	ensure proper ``floating'' placement of tables, use the environment
	\textbf{figure*} to enclose the figure and its caption.
	and don't forget to end the environment with
	{figure*}, not {figure}!

	\begin{figure*}
		\centering
		%\includegraphics{flies.eps}
		\caption{A sample black and white graphic
			that needs to span two columns of text.}
	\end{figure*}

	\subsection{Theorem-like Constructs}
	Other common constructs that may occur in your article are
	the forms for logical constructs like theorems, axioms,
	corollaries and proofs.  There are
	two forms, one produced by the
	command \texttt{{\char'134}newtheorem} and the
	other by the command \texttt{{\char'134}newdef}; perhaps
	the clearest and easiest way to distinguish them is
	to compare the two in the output of this sample document:

	This uses the \textbf{theorem} environment, created by
	the\linebreak\texttt{{\char'134}newtheorem} command:
	\newtheorem{theorem}{Theorem}
	\begin{theorem}
		Let $f$ be continuous on $[a,b]$.  If $G$ is
		an antiderivative for $f$ on $[a,b]$, then
		\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
	\end{theorem}

	The other uses the \textbf{definition} environment, created
	by the \texttt{{\char'134}newdef} command:
	\newdef{definition}{Definition}
	\begin{definition}
		If $z$ is irrational, then by $e^z$ we mean the
		unique number which has
		logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
	\end{definition}

	Two lists of constructs that use one of these
	forms is given in the
	\textit{Author's  Guidelines}.

	There is one other similar construct environment, which is
	already set up
	for you; i.e. you must \textit{not} use
	a \texttt{{\char'134}newdef} command to
	create it: the \textbf{proof} environment.  Here
	is a example of its use:
	\begin{proof}
		Suppose on the contrary there exists a real number $L$ such that
		\begin{displaymath}
		\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
		\end{displaymath}
		Then
		\begin{displaymath}
		l=\lim_{x\rightarrow c} f(x)
		= \lim_{x\rightarrow c}
		\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
		= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
		\frac{f(x)}{g(x)} = 0\cdot L = 0,
		\end{displaymath}
		which contradicts our assumption that $l\neq 0$.
	\end{proof}

	Complete rules about using these environments and using the
	two different creation commands are in the
	\textit{Author's Guide}; please consult it for more
	detailed instructions.  If you need to use another construct,
	not listed therein, which you want to have the same
	formatting as the Theorem
	or the Definition\cite{salas:calculus} shown above,
	use the \texttt{{\char'134}newtheorem} or the
	\texttt{{\char'134}newdef} command,
	respectively, to create it.

	\subsection*{A {\secit Caveat} for the \TeX\ Expert}
	Because you have just been given permission to
	use the \texttt{{\char'134}newdef} command to create a
	new form, you might think you can
	use \TeX's \texttt{{\char'134}def} to create a
	new command: \textit{Please refrain from doing this!}
	Remember that your \LaTeX\ source code is primarily intended
	to create camera-ready copy, but may be converted
	to other forms -- e.g. HTML. If you inadvertently omit
	some or all of the \texttt{{\char'134}def}s recompilation will
	be, to say the least, problematic.

	\section{Conclusions}
	This paragraph will end the body of this sample document.
	Remember that you might still have Acknowledgments or
	Appendices; brief samples of these
	follow.  There is still the Bibliography to deal with; and
	we will make a disclaimer about that here: with the exception
	of the reference to the \LaTeX\ book, the citations in
	this paper are to articles which have nothing to
	do with the present subject and are used as
	examples only.
	%\end{document}  % This is where a 'short' article might terminate

	%ACKNOWLEDGMENTS are optional
	\section{Acknowledgments}
	This section is optional; it is a location for you
	to acknowledge grants, funding, editing assistance and
	what have you.  In the present case, for example, the
	authors would like to thank Gerald Murray of ACM for
	his help in codifying this \textit{Author's Guide}
	and the \textbf{.cls} and \textbf{.tex} files that it describes.

	%
	% The following two commands are all you need in the
	% initial runs of your .tex file to
	% produce the bibliography for the citations in your paper.
	\bibliographystyle{abbrv}
	\bibliography{biblio}  
	% You must have a proper ".bib" file
	%  and remember to run:
	% latex bibtex latex latex
	% to resolve all references
	%
	% ACM needs 'a single self-contained file'!
	%
	%APPENDICES are optional
	%\balancecolumns
	\appendix
	%Appendix A
	\section{Headings in Appendices}
	The rules about hierarchical headings discussed above for
	the body of the article are different in the appendices.
	In the \textbf{appendix} environment, the command
	\textbf{section} is used to
	indicate the start of each Appendix, with alphabetic order
	designation (i.e. the first is A, the second B, etc.) and
	a title (if you include one).  So, if you need
	hierarchical structure
	\textit{within} an Appendix, start with \textbf{subsection} as the
	highest level. Here is an outline of the body of this
	document in Appendix-appropriate form:
	\subsection{Introduction}
	\subsection{The Body of the Paper}
	\subsubsection{Type Changes and  Special Characters}
	\subsubsection{Math Equations}
	\paragraph{Inline (In-text) Equations}
	\paragraph{Display Equations}
	\subsubsection{Citations}
	\subsubsection{Tables}
	\subsubsection{Figures}
	\subsubsection{Theorem-like Constructs}
	\subsubsection*{A Caveat for the \TeX\ Expert}
	\subsection{Conclusions}
	\subsection{Acknowledgments}
	\subsection{Additional Authors}
	This section is inserted by \LaTeX; you do not insert it.
	You just add the names and information in the
	\texttt{{\char'134}additionalauthors} command at the start
	of the document.
	\subsection{References}
	Generated by bibtex from your ~.bib file.  Run latex,
	then bibtex, then latex twice (to resolve references)
	to create the ~.bbl file.  Insert that ~.bbl file into
	the .tex source file and comment out
	the command \texttt{{\char'134}thebibliography}.
	% This next section command marks the start of
	% Appendix B, and does not continue the present hierarchy
	\section{More Help for the Hardy}
	The sig-alternate.cls file itself is chock-full of succinct
	and helpful comments.  If you consider yourself a moderately
	experienced to expert user of \LaTeX, you may find reading
	it useful but please remember not to change it.
	%\balancecolumns % GM June 2007
	% That's all folks!
\end{document}