### 2016-05-11 exp_setup.tex
chap4 added

\section{Evaluation}
This chapter describes the experimental setup for our project including the servers used, the software and the platforms involved, the languages used, and the task queue system that was used for parallelism. We follow this up with our evaluation of the system, with a test suite, and proof of concept examples.

\input{Experiment/sys_config}
\input{Experiment/platform}
\input{Experiment/languages}
\input{Experiment/celery}
\input{System/test_plan}
\input{System/proof_of_concept}