### 2016-05-13 Results/data_analysis.tex

\begin{table}[!htbp]
	\centering
	\begin{tabular}{|c|l|c|}
		\hline
		\multicolumn{1}{|c|}{\textbf{S.No}} &
		\multicolumn{1}{c|}{\textbf{Type of Injection}} &
		\multicolumn{1}{p{3cm}|}{\centering \textbf{No. of e-mails received}}\\
		\hline
		1 & E-Mail Header Injections with \texttt{bcc} header & 265\\
		\hline
		2 & E-Mail Header Injections with \texttt{x-check} header & 214\\
		\hline
		3 & \texttt{To header} injections alone & 142\\
		\hline
		4 & E-Mail Header Injections with \texttt{bcc} and \texttt{x-check} headers & 188\\
		\hline
		5 & Both \texttt{To header} injections and x-check headers &
		11\\
		\hline
		6 & \texttt{x-check} headers found in \texttt{nuser} e-mails & 56\\
		\hline
		7 & Unique \texttt{x-check} headers found in \texttt{nuser} e-mails & 26\\
		\hline
		8 & Total successful injections (1 + 3 + 7) & 433\\

		\hline
	\end{tabular}
	\caption[\titlecap{Analysis of the data}]{Classification of the e-mails that we received into broad categories of the vulnerability.}
	\label{tab:analysis}
\end{table}