### 2016-05-13 Background/py.tex
fixing code samples

\lstset{language=Python,caption={Python program with e-mail header injection vulnerability.},label={code:pyemi}}
\begin{lstlisting}
from email.parser import Parser
import cgi
form = cgi.FieldStorage()
to = form["email"] # input() exhibits the same behavior
msg = """To: """ + to + """\n
From: <user@example.com>\n
Subject: Test message\n\n
Body would go here\n"""

f = FeedParser() # Parser.parsestr() also 
# contains the same vulnerability
f.feed(msg)
headers = FeedParser.close(f)

# attack string => 'sai@sai.com\nBCC:spc@spc.com'
# for form["email"]

# to:sai@sai.com AND bcc:spc@spc.com 
# are added to the headers
print 'To: %s' % headers['to']
print 'BCC: %s' % headers['bcc']
\end{lstlisting}