### 2016-05-14 exp_setup.tex

\section{Implementation}

\input{Experiment/sys_config}
%\input{Experiment/platform}
%\input{Experiment/languages}
%\input{Experiment/celery}
%\input{System/test_plan}
\input{System/proof_of_concept}