### 2016-05-17 paper.tex
Removing cruft from the files. Adding makefile and .gitignore

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate-05-2015}

\usepackage[square,numbers]{natbib}
\usepackage[justification=centering]{caption}
\usepackage[pageanchor=true,plainpages=false,pdfpagelabels,bookmarks,bookmarksnumbered,hidelinks]{hyperref}
\usepackage{color}
\definecolor{mygray}{rgb}{0.6,0.6,0.6}
\definecolor{lightgray}{rgb}{0.92,0.92,0.92}
\definecolor{darkgreen}{rgb}{0,0.7,0}

% for footnotes in tables
\usepackage{tablefootnote}

% for title caps
\usepackage{titlecaps}
\Addlcwords{and, the, or, of, that, our, by, a, prevent, for}

% default options for listings
\usepackage{listings}
\usepackage{listingsutf8}
\usepackage{textcomp}     % access \textquotesingle
\lstset{
	backgroundcolor=\color{lightgray},
	basicstyle=\small,
	breaklines=true, 
	captionpos=b,
	commentstyle=\color{darkgreen}, 
	frame=single,
	keywordstyle=\color{blue}, 
	numbers=left,
	numbersep=5pt,
	numberstyle=\tiny\color{mygray},
	rulecolor=\color{black},
	showstringspaces=false,
	upquote=true
}

\newcommand{\dq}[1]{``{#1}''}
%% THIS IS THE DATA FOR OUR THESIS, UPDATING HERE WILL UPDATE EVERYWHERE %%
\newcommand{\urls}{21,675,680}

\newcommand{\forms}{6,794,917}
\newcommand{\formsDelta}{31.35\%}

\newcommand{\emailforms}{1,132,157}
\newcommand{\emailformsDelta}{16.66\%}

\newcommand{\fuzzed}{934,016}
\newcommand{\fuzzedDelta}{82.50\%}

\newcommand{\recd}{52,724}
\newcommand{\recdDelta}{5.64\%}

\newcommand{\malfuzzed}{46,156}
\newcommand{\malfuzzedDelta}{87.54\%}

\newcommand{\success}{496}
\newcommand{\successDelta}{1.07\%}

\newcommand{\domains}{222}
\newcommand{\ips}{292}
% these refer to unique domains, not unique forms
\newcommand{\uniqueforms}{1,019,921}
\newcommand{\uniqueemailforms}{197,570}

\begin{document}

	% Copyright
	\setcopyright{acmcopyright}
	%\setcopyright{acmlicensed}
	%\setcopyright{rightsretained}
	%\setcopyright{usgov}
	%\setcopyright{usgovmixed}
	%\setcopyright{cagov}
	%\setcopyright{cagovmixed}

	% DOI
	%\doi{10.475/123_4}

	% ISBN
	%\isbn{123-4567-24-567/08/06}

	%Conference
	%\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

	%\acmPrice{\$15.00}

	%
	% --- Author Metadata here ---
	%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
	%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
	%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
	% --- End of Author Metadata ---

	%\title{E-jection fraction - Tracking how your website pumps out
	%E-Mails}
    \title{The Prevalence of E-Mail Header Injections on the World
      Wide Web}
	%
	% You need the command \numberofauthors to handle the 'placement
	% and alignment' of the authors beneath the title.
	%
	% For aesthetic reasons, we recommend 'three authors at a time'
	% i.e. three 'name/affiliation blocks' be placed beneath the title.
	%
	% NOTE: You are NOT restricted in how many 'rows' of
	% "name/affiliations" may appear. We just ask that you restrict
	% the number of 'columns' to three.
	%
	% Because of the available 'opening page real-estate'
	% we ask you to refrain from putting more than six authors
	% (two rows with three columns) beneath the article title.
	% More than six makes the first-page appear very cluttered indeed.
	%
	% Use the \alignauthor commands to handle the names
	% and affiliations for an 'aesthetic maximum' of six authors.
	% Add names, affiliations, addresses for
	% the seventh etc. author(s) as the argument for the
	% \additionalauthors command.
	% These 'additional authors' will be output/set for you
	% without further effort on your part as the last section in
	% the body of your article BEFORE References or any Appendices.

	\numberofauthors{1} %  in this sample file, there are a *total*
	% of EIGHT authors. SIX appear on the 'first-page' (for formatting
	% reasons) and the remaining two appear in the \additionalauthors section.
	%

    % Note: For CCS submission we need to have it anonymous

	\author{
		% You can go ahead and credit any number of authors here,
		% e.g. one 'row of three' or two rows (consisting of one row of three
		% and a second row of one, two or three).
		%
		% The command \alignauthor (no curly braces needed) should
		% precede each author name, affiliation/snail-mail address and
		% e-mail address. Additionally, tag each line of
		% affiliation/address with \affaddr, and tag the
		% e-mail address with \email.
		%
		% 1st. author
		%% \alignauthor
		%% Sai Prashanth Chandramouli\\
		%% \affaddr{Arizona State University}\\
		%% \email{saipc@asu.edu}
		%% % 2nd. author
		%% \alignauthor
		%% Adam Doup\'{e}\titlenote{The secretary disavows
		%% 	any knowledge of this author's actions.}\\
		%% \affaddr{Arizona State University}\\
		%% \email{doupe@asu.edu}
		%% % 3rd. author
		%% \alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
		%% 	one who did all the really hard work.}\\
		%% \affaddr{The Th{\o}rv{\"a}ld Group}\\
		%% \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
		%% \affaddr{Hekla, Iceland}\\
		%% \email{larst@affiliation.org}
		%% \and  % use '\and' if you need 'another row' of author names
		%% % 4th. author
		%% \alignauthor Lawrence P. Leipuner\\
		%% \affaddr{Brookhaven Laboratories}\\
		%% \affaddr{Brookhaven National Lab}\\
		%% \affaddr{P.O. Box 5000}\\
		%% \email{lleipuner@researchlabs.org}
		%% % 5th. author
		%% \alignauthor Sean Fogarty\\
		%% \affaddr{NASA Ames Research Center}\\
		%% \affaddr{Moffett Field}\\
		%% \affaddr{California 94035}\\
		%% \email{fogartys@amesres.org}
		%% % 6th. author
		%% \alignauthor Charles Palmer\\
		%% \affaddr{Palmer Research Laboratories}\\
		%% \affaddr{8600 Datapoint Drive}\\
		%% \affaddr{San Antonio, Texas 78229}\\
		%% \email{cpalmer@prl.com}
	}
	% There's nothing stopping you putting the seventh, eighth, etc.
	% author on the opening page (as the 'third row') but we ask,
	% for aesthetic reasons that you place these 'additional authors'
	% in the \additional authors block, viz.
	%% \additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
	%% 	email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
	%% 	(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
	%% \date{30 July 1999}
	% Just remember to make sure that the TOTAL number of authors
	% is the number that will appear on the first page PLUS the
	% number that will appear in the \additionalauthors section.

	\maketitle
	\input{abstract}

	%
	% The code below should be generated by the tool at
	% http://dl.acm.org/ccs.cfm
	% Please copy and paste the code instead of the example below. 
	%
	%% \begin{CCSXML}
	%% 	<ccs2012>
	%% 	<concept>
	%% 	<concept_id>10010520.10010553.10010562</concept_id>
	%% 	<concept_desc>Computer systems organization~Embedded systems</concept_desc>
	%% 	<concept_significance>500</concept_significance>
	%% 	</concept>
	%% 	<concept>
	%% 	<concept_id>10010520.10010575.10010755</concept_id>
	%% 	<concept_desc>Computer systems organization~Redundancy</concept_desc>
	%% 	<concept_significance>300</concept_significance>
	%% 	</concept>
	%% 	<concept>
	%% 	<concept_id>10010520.10010553.10010554</concept_id>
	%% 	<concept_desc>Computer systems organization~Robotics</concept_desc>
	%% 	<concept_significance>100</concept_significance>
	%% 	</concept>
	%% 	<concept>
	%% 	<concept_id>10003033.10003083.10003095</concept_id>
	%% 	<concept_desc>Networks~Network reliability</concept_desc>
	%% 	<concept_significance>100</concept_significance>
	%% 	</concept>
	%% 	</ccs2012>  
	%% \end{CCSXML}

	%% \ccsdesc[500]{Computer systems organization~Embedded systems}
	%% \ccsdesc[300]{Computer systems organization~Redundancy}
	%% \ccsdesc{Computer systems organization~Robotics}
	%% \ccsdesc[100]{Networks~Network reliability}

	%
	% End generated code
	%

	%
	%  Use this command to print the description
	%
	% \printccsdesc

	% We no longer use \terms command
	%\terms{Theory}

	% \keywords{ACM proceedings; \LaTeX; text tagging}

	\input{introduction}

	\input{problem_background}

	\input{sys_design}

	\input{exp_setup}

	\input{results}

	\input{discussion}

	\input{rel_work}

	\input{conclusion}

	%
	% The following two commands are all you need in the
	% initial runs of your .tex file to
	% produce the bibliography for the citations in your paper.
	\bibliographystyle{abbrv}
	\bibliography{biblio}  
	% You must have a proper ".bib" file
	%  and remember to run:
	% latex bibtex latex latex
	% to resolve all references
	%
	%\balancecolumns
\end{document}