### 2016-05-19 exp_setup.tex
Cleanup, removed unused files

\section{Implementation}

\input{Experiment/sys_config}
\input{System/proof_of_concept}