### 2016-05-19 Background/php.tex
rewrote code samples to anonymize the paper, removed references to myself and to asu

\lstset{language=PHP,caption={PHP program with e-mail header injection vulnerability.},label={code:phpemi}}
\begin{lstlisting}
$from = $_REQUEST['email'];
$subject = 'Hello XYZ';
$message = 'We need you to reset your password';
$to = 'xyz@xyz.edu';

// example attack string to be 
// injected as the value for
// $_REQUEST['email'] => 
// 's@s.com\nCC:spc@spc.com'
$retValue = mail($to, $subject, $message, "From: $from");
// E-Mail gets sent to both 
// xyz@xyz.edu AND spc@spc.com
\end{lstlisting}