### 2016-05-23 Background/php.tex

\begin{lstlisting}[language=PHP,caption={PHP program with e-mail
      header injection vulnerability.},label={code:phpemi}, float]
$from = $_REQUEST['email'];
$subject = 'Hello XYZ';
$message = 'We need you to reset your password';
$to = 'xyz@example.com';

// example attack string to be 
// injected as the value for
// $_REQUEST['email'] => 
// 'abc@example.com\nCC:spc@example.com'
$retValue = mail($to, $subject, $message, "From: $from");
// E-Mail gets sent to both 
// xyz@example.com AND spc@example.com
\end{lstlisting}