### 2016-05-24 Results/data_analysis.tex

\begin{table*}[tbp]
	\centering
	\begin{tabular}{|l|c|}
		\hline
		\multicolumn{1}{|c|}{\textbf{Type of Injection}} &
		\multicolumn{1}{p{3cm}|}{\centering \textbf{No. of e-mails received}}\\
		\hline
		E-Mail Header Injections with \texttt{bcc} header & \ehibcc \\
		\hline
		E-Mail Header Injections with \texttt{x-check} header & \ehixcheck \\
		\hline
		\texttt{To header} injections alone & \ehito \\
		\hline
		E-Mail Header Injections with \texttt{bcc} and \texttt{x-check} headers & \ehibccxcheck \\
		\hline
		Both \texttt{To header} injections and x-check headers &
		\ehitoxcheck \\
		\hline
		\texttt{x-check} headers found in \texttt{nuser} e-mails & \ehinuserxcheck \\
		\hline
		Unique \texttt{x-check} headers found in \texttt{nuser} e-mails & \ehiuniquenuserxcheck \\
		\hline
		Total successful injections (1 + 3 + 7) & \success \\

		\hline
	\end{tabular}
	\caption[\titlecap{Analysis of the data}]{Classification of the e-mails that we received into broad categories of the vulnerability.}
	\label{tab:analysis}
\end{table*}