### 2016-05-24 Results/pre_result.tex

\subsection{Collected Data}

% TODO: Need to put in a number here, and discuss the high level
% crawl. - DONE

From our extensive crawl of the web, we were able to gather the data
shown in Table~\ref{tab:data}. We ran the system for 76 days, during which our system crawled \urls unique URLs,
and found a total of \forms\ forms from \uniqueforms\ unique domains. Out of these forms, our system
found \emailforms\ forms that contained an \email field, from \uniqueemailforms\ unique domains.

\input{data}