### 2016-05-24 Discussion/ethics.tex
fixed related work, starting ethics

\subsection{Ethics}
To make sure that our system did not cause any major harm to the websites that we crawled, we made sure that we did not inject any special characters other than the newline character. These payl(like SQL Injection and Cross-Site Scripting)