### 2016-05-24 Discussion/ethics.tex
added ethics section

\subsection{Ethics}
To make sure that our system did not cause any harm to the websites that we crawled, we made sure that we did not inject any special characters other than the newline character% (characters that a developer would assume an average user could put in)
. We also had an information website\footnote{Removed for the sake of anonymity} that described what \ehi was, and contained our contact details in case the developers of the websites we crawled wanted to contact us. We maintained a separate blacklist of websites that the owners did not want us to crawl, and ensured that our system did not crawl their websites.