### 2016-05-24 Discussion/ethics.tex
Changed websites to web applications

\subsection{Ethics}

To make sure that our system did not cause any harm to the web applications that we crawled, we made sure that we did not inject any special characters other than the newline character% (characters that a developer would assume an average user could put in)
. We also had an information website at the IP that we crawled from\footnote{Removed for the sake of anonymity} that described what \ehi was, and contained our contact details in case the developers of the web applications we crawled wanted to contact us. We maintained a separate blacklist of domain names that the owners did not want us to crawl, and ensured that our system did not crawl their domains.