### 2016-10-24 Results/virus_data.tex

\subsection{Emails with Malicious Attachments}
We checked the \emails received by the `reguser' account (which are injected with regular \email addresses and not malicious data) for the presence of attachments that may contain malicious software, which will indicate that the server itself is not benign.

To do this, we passed the \totalattachmentcount attachments we received to VirusTotal~\cite{virustotal} -- an online virus and malware scanner that checks for the presence of malware by running 40+ virus scanners on the uploaded files. We found that out of the \totalattachmentcount attachments, \totalvirusemails \emails contained malicious attachments, out of which \totalvirusattachmentcount were from unique domains. 
%It is to be noted that none of these domains were found on the spam analysis we did earlier, indicating that if these could be exploited, we could send emails with malicious attachments to other people using these servers.