### 2016-10-24 exp_setup.tex

\section{Validation}

%\input{Experiment/sys_config}
\input{System/proof_of_concept}