### 2016-10-24 System/approach.tex

\subsection{Approach}
\label{sys:appr}
We took a black-box approach to measure the prevalence of \ehi vulnerability on the web. Black-box testing~\cite{Beizer:1995:BTT:202699} is a way to examine the functionality of an application without analyzing its source code.

Black-box testing allows our system to detect \ehi vulnerabilities in \emph{any} server-side language (not simply those we identified in Section~\ref{languages}). The overall architecture of our system is presented in Figure~\ref{fig:overall}. The individual components are discussed below.

\begin{figure*}
	\centering
	\includegraphics[width=.75\textwidth, height=.25\textheight]{overall}
	\caption{Overall system architecture.}
	\label{fig:overall}
\end{figure*}