### 2016-10-24 Background/py.tex

\begin{lstlisting}[language=Python,caption={Python program with e-mail
      header injection vulnerability.},label={code:pyemi}, float]
from email.parser import Parser
import cgi
form = cgi.FieldStorage()
to = form["email"]
msg = """To: """ + to + """\n
From: <user@example.com>\n
Subject: Test message\n\n
Body would go here\n"""

f = FeedParser() 
f.feed(msg)
headers = FeedParser.close(f)
print 'To: %s' % headers['to']
print 'BCC: %s' % headers['bcc']
\end{lstlisting}