### 2016-10-24 Background/ruby.tex

\begin{lstlisting}[language=Ruby,caption={Ruby program with e-mail
      header injection vulnerability.},label={code:rubyemi}, float]
require 'sinatra'
require 'net/smtp'
get '/hello' do
email = params[:email]
message = """
    From: Sch <sch@example.com>
    Subject: SMTP e-mail test
    To: #{email}

    This is a test e-mail message.
"""
Net::SMTP.start('localhost',1025) do |smtp|
smtp.send_message message, 'sch@example.com', 'to@todomain.com'
end
end
\end{lstlisting}