### 2016-10-24 Results/data_analysis.tex

\begin{table}[tbp]
	\centering
	\scriptsize
	\begin{tabular}{|p{.32\textwidth}|p{.13\textwidth}|}
		\hline
		\textbf{Type of Injection} & \textbf{No. of e-mails received}\\
		\hline
		E-Mail Header Injections with \texttt{bcc} header & \ehibcc \\
		\hline
		E-Mail Header Injections with \texttt{x-check} header & \ehixcheck \\
		\hline
		\texttt{To header} injections alone & \ehito \\
		\hline
		E-Mail Header Injections with \texttt{bcc} and \texttt{x-check} headers & \ehibccxcheck \\
		\hline
		Both \texttt{To header} injections and x-check headers &
		\ehitoxcheck \\
		\hline
		\texttt{x-check} headers found in \texttt{nuser} e-mails & \ehinuserxcheck \\
		\hline
		Unique \texttt{x-check} headers found in \texttt{nuser} e-mails & \ehiuniquenuserxcheck \\
		\hline
		Total successful injections (1 + 3 + 7) & \success \\

		\hline
	\end{tabular}
	\caption[\titlecap{Analysis of the data}]{Classification of the e-mails that we received into broad categories of the vulnerability.}
	\label{tab:analysis}
\end{table}