### 2016-10-24 Results/fuzz_result.tex

%\subsection{Fuzzed Data and Received \Emails}

Table~\ref{tab:fuzzed_data} shows the quantity of \emails that we received for the benign and malicious payloads. 
\input{Results/fuzzed_data}

\noindent\textbf{\Email received from forms.} The \emails that we
received can be categorized into two categories. (1) \Emails due to
regular payload: This represents the total number of web applications
that sent \emails to us. This indicates that we were able to
successfully submit the forms on these sites to trigger the web
application to send an \email. (2) \Emails due to malicious payload:
Once we receive an \email from a web application due to the regular
payload, we fuzz those forms with the malicious payloads. This field
represents the total number of unique URLs that are contain an \ehi
vulnerability.