### 2016-10-24 System/sys_arch.tex

\subsection{System Architecture}
\label{sys:arch}
Our system can be broadly divided into two modules: Data Gathering and Payload Injection.

The Data Gathering module is responsible for the following activities:
(1) Interface with the Crawler (Section~\ref{Comp:Crawler}) and
receive the URLs, (2) Parse the HTML for the corresponding URL and
store the relevant form data (Section~\ref{Comp:FP}), and (3) Check
for the presence of HTML forms that could allow a user to send/receive
\email and store references to these forms (Section~\ref{Comp:EMFC}).

The Payload Injection module is responsible for the following
activities: (1) Retrieve the HTML forms that could allow a user to
send/receive \email and reconstruct these forms
(Section~\ref{Comp:EMFR}), (2) Inject these forms with benign data
(non-malicious payloads) and generate an HTTP request to the
corresponding URL (Section~\ref{Comp:Fuzzer}), (3) Analyze the
received \emails, extracting the \email header fields and checking for
the presence of the injected payloads (Section~\ref{Comp:EMA}), and
(4) Inject the HTTP requests that sent us \emails with malicious
payloads, and generate an HTTP request to the corresponding URL to
check if an \ehi vulnerability exists in that web application
(Section~\ref{Comp:Fuzzer}).

	%% The functionality of each component is discussed further in the `Components' section (Section~\ref{Comp}). The Payload Injection pipeline is not a linear, but cyclic process, as we inject different payloads and analyze the received \emails.