### 2016-10-25 System/approach.tex

\subsection{Approach}
\label{sys:appr}
We took a black-box approach to measure the prevalence of \ehi
vulnerability on the web. Black-box
testing~\cite{Beizer:1995:BTT:202699} is a way to examine the
functionality of an application without analyzing its source code. Black-box testing allows our system to detect \ehi vulnerabilities in \emph{any} server-side language (not simply those we identified in Section~\ref{languages}). The overall architecture of our system is presented in Figure~\ref{fig:overall}.

\begin{figure}
	\centering
	\includegraphics[width=.5\textwidth]{overall_crop}
	\caption{Overall system architecture.}
    \vspace{-5ex}
	\label{fig:overall}
\end{figure}