### 2016-10-25 Results/fuzz_result.tex
need to cut down 13 lines more woohoo

%Table~\ref{tab:fuzzed_data} shows the quantity of \emails that we received for the benign and malicious payloads. 
%\input{Results/fuzzed_data}
%
%\noindent\textbf{\Email received from forms.} The \emails that we
%received can be categorized into two categories. (1) \Emails due to
%regular payload: This represents the total number of web applications
%that sent \emails to us. This indicates that we were able to
%successfully submit the forms on these sites to trigger the web
%application to send an \email. (2) \Emails due to malicious payload:
%Once we receive an \email from a web application due to the regular
%payload, we fuzz those forms with the malicious payloads. This field
%represents the total number of unique URLs that are contain an \ehi
%vulnerability.